Versioned data: why it is needed and how it can be achieved (easily and cheaply)

### 2017-09-21 ms.tex
comment out old abstract. address #21

% I think that the real issue with the first approach is that it's not
% straightforward for a small group to set up; that is probably packed
% away in "unsustainable" but could be repgrased to make it clearer
% perhaps?  Breaking down barriers is always a good aim
%
% I really think that something like CKAN (though it doesn't support
% holding data if I remember correctly) could totally be a complete
% solution to the sorts of issues that we have here.
%
% The "changing over time" is introduced as a problem associated with
% medium sized data but I think it comes from nowhere in the current
% draft.  It's more a separate concern than the big/small/medium size
% angle.
The sharing and re-use of data has become a cornerstone of modern science. To date, the curation and distribution of data has largely been conducted under one of two frameworks: data is either deposited in large, collaborative databases (often run by institutions and staffed by computational professionals) or else published as static files alongside a specific publication or as a standalone product released by a research group. For most medium-sized datasets, the former is overkill and likely unwieldy and unsustainable, and the latter -- while convenient -- is rather limited as it does not allow the dataset to improve nor to grow over time. This indicates a substantial conceptual and technological gap in how we curate and share data. In this article, we argue that adopting the paradigm and processes for versioned data, analogous to software versioning, can help fill this gap. We introduce a system called Versioned Data Delivery (VDD) and present novel computational tools for creating, archiving, and distributing versioned data easily, quickly, and cheaply. These new tools allow for individual research groups to shift from a static model of data curation to a dynamic and versioned model that naturally matches the scientific process.

>>>>

% I think that the real issue with the first approach is that it's not
% straightforward for a small group to set up; that is probably packed
% away in "unsustainable" but could be repgrased to make it clearer
% perhaps?  Breaking down barriers is always a good aim
%
% I really think that something like CKAN (though it doesn't support
% holding data if I remember correctly) could totally be a complete
% solution to the sorts of issues that we have here.
%
% The "changing over time" is introduced as a problem associated with
% medium sized data but I think it comes from nowhere in the current
% draft.  It's more a separate concern than the big/small/medium size
% angle.
% The sharing and re-use of data has become a cornerstone of modern science. To date, the curation and distribution of data has largely been conducted under one of two frameworks: data is either deposited in large, collaborative databases (often run by institutions and staffed by computational professionals) or else published as static files alongside a specific publication or as a standalone product released by a research group. For most medium-sized datasets, the former is overkill and likely unwieldy and unsustainable, and the latter -- while convenient -- is rather limited as it does not allow the dataset to improve nor to grow over time. This indicates a substantial conceptual and technological gap in how we curate and share data. In this article, we argue that adopting the paradigm and processes for versioned data, analogous to software versioning, can help fill this gap. We introduce a system called Versioned Data Delivery (VDD) and present novel computational tools for creating, archiving, and distributing versioned data easily, quickly, and cheaply. These new tools allow for individual research groups to shift from a static model of data curation to a dynamic and versioned model that naturally matches the scientific process.
